Learning from Massive Noisy Labeled Data for Image Classification

### 2015-04-13 paper.tex
Update exp and conclusion

Our goal is to build an end-to-end deep learning system that is capable of training with both limited clean labels and massive noisy labels more effectively. Fig.~\ref{fig:overview} shows the framework of our approach. We collect $1,000,000$ clothing images from online shopping websites. Each image is automatically assigned with a noisy label according to the keywords in its surrounding text. We manually refine $72,409$ image labels, which constitute a clean sub-dataset. All the data are then used to train CNNs, while the major challenge is to identify and correct wrong labels during the training process.

To cope with this challenge, we extend CNNs with a novel probabilistic model, which infers the true labels and uses them to supervise the training of the network. Our work is inspired by~\cite{sukhbaatar2014learning}, which modified a CNN by inserting a linear layer on top of the softmax layer to map clean labels to noisy labels. However,~\cite{sukhbaatar2014learning} assumed noisy labels are conditionally independent with input images given clean labels. By examining our collected dataset, we find that this assumption is too strong to fit real-world data well. For example, in Fig.~\ref{fig:noise_cluster}, all the images should belong to ``Hoodie''. The top five are correct while the bottom five are either mislabeled as ``Windbreaker'' or ``Jacket''. Since different sellers have their own bias on different categories, they may provide wrong keywords for similar clothes. We can capture these visual patterns to estimate how likely an image is mislabeled. Based on these observations, we further introduce two types of label noise:
\begin{itemize}
\item \textbf{Confusing noise} makes the noisy label reasonably wrong. It usually occurs when the image content is confusing (\eg, the samples with ``?'' in Fig.~\ref{fig:overview}).
\item \textbf{Pure random noise} makes the noisy label totally wrong. It is often caused by either the mismatch between an image and its surrounding text, or the false conversion from the text to label (\eg, the samples with ``$\times$'' in Fig.~\ref{fig:overview}).
\end{itemize}

@@

\section{Label Noise Model} % (fold)
\label{sec:label_noise_model}
We target on learning a classifier from a set of images with noisy labels. To be specific, we have a noisy labeled dataset $\mathcal{D}_{\eta}=\left\{\left(\vec{x}^{(1)}, \tilde{y}^{(1)}\right), \dots, \left(\vec{x}^{(N)}, \tilde{y}^{(N)}\right)\right\}$ with $n$-th image $\vec{x}^{(n)}$ and its corresponding noisy label $\tilde{y}^{(n)} \in \{1,\dots,L\}$, where $L$ is the number of classes. We describe how the noisy label is generated by using a probabilistic graphical model shown in Fig.~\ref{fig:noise_pgm}.

@@

Following this assignment rule, we define the conditional probability of the noisy label as
\begin{equation} \label{eq:likelihood}
    p(\tilde{\vec{y}}|\vec{y},\vec{z}) = \begin{cases}
        \tilde{\vec{y}}^{T}\mat{I}\vec{y} & \text{if } \vec{z}_1 = 1 \\
        \frac{1}{L-1}\tilde{\vec{y}}^{T}(\mat{U} - \mat{I})\vec{y} & \text{if } \vec{z}_2 = 1 \\
        \tilde{\vec{y}}^{T}\mat{C}\vec{y} & \text{if } \vec{z}_3 = 1,
    \end{cases}
\end{equation}
where $\mat{I}$ is the identity matrix, $\mat{U}$ is the unit matrix (all the elements are ones), $\mat{C}$ is a sparse stochastic matrix with $\mathrm{tr}(\mat{C})=0$ and $\mat{C}_{ij}$ denotes the confusion probability between classes $i$ and $j$. Then we can derive from Fig.~\ref{fig:noise_pgm} the joint distribution of $\tilde{\vec{y}}, \vec{y}$ and $\vec{z}$ conditioning on $\vec{x}$,
\begin{equation} \label{eq:joint_distribution}
  p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}) = p(\tilde{\vec{y}} | \vec{y}, \vec{z}) p(\vec{y} | \vec{x}) p(\vec{z} | \vec{x}).
\end{equation}

While the class label probability distribution $p(\vec{y} | \vec{x})$ is comprehensible, the semantic meaning of $p(\vec{z} | \vec{x})$ needs extra clarification: it represents how confusing the image content is. Specific to our clothing classification problem, $p(\vec{z} | \vec{x})$ can be affected by different factors, including background clutter, image resolution, the style and material of the clothes. Some examples are shown in Fig.~\ref{fig:compare_noise_level}.

@@

We integrate the proposed label noise model into a deep learning framework. As demonstrated in Fig.~\ref{fig:framework}, we predict the probability $p(\vec{y}|\vec{x})$ and $p(\vec{z}|\vec{x})$ by using two independent CNNs. Moreover, we append a label noise model layer at the end, which takes as input the CNNs' prediction scores and the observed noisy label. Stochastic Gradient Ascent with backpropagation technique is used to approximately optimize the whole network. In each forward pass, the label noise model layer computes the posterior of latent variables according to Eq.~\eqref{eq:posterior}. While in the backward pass, it computes the gradients according to Eq.~\eqref{eq:gradient}.

@@

After both CNN components are properly pretrained, we can start to train the whole network with massive noisy labeled data. However, some practical issues need further discussion. First, if we merely use noisy labels, we will lose precious knowledge that we have gained before and the model could be drifted. Therefore, we need to mix the data with clean labels into our training set, which is depicted in Fig.~\ref{fig:framework} as the extra supervisions for the two CNNs. Then each CNN receives two kinds of gradients, one is from the clean labels and the other is from the noisy labels. We denote them by $\Delta_c$ and $\Delta_n$, respectively. A potential problem is that $|\Delta_c| \ll |\Delta_n|$, because clean data is much less than noisy data. To deal with this problem, we bootstrap the clean data $\mathcal{D}_c$ to half amount of the noisy data $\mathcal{D}_{\eta}$. In our experiments, we find that the performance of the classifier drops significantly without upsampling, but it is not sensitive with the upsampling ratio as long as the number of clean and noisy samples are in the same order.

@@

The confusion matrix between clean and noisy labels is presented in Fig.~\ref{fig:cm_strong_weak}. We can see that the overall accuracy is $61.54\%$, and some pairs of classes are very confusing with each other (\eg Knitwear and Sweater), which means that the noisy labels are not so reliable.

@@

\subsection{Evaluation on the Collected Dataset} % (fold)
\label{sub:evaluation}
We validate our method through a series of experiments conducted on the collected dataset. Our implementation is based on Caffe~\cite{jia2014caffe}, and the \texttt{bvlc\_reference\_caffenet}\footnote{\url{http://caffe.berkeleyvision.org/model_zoo.html}} is chosen as the baseline model, which approximates AlexNet~\cite{krizhevsky2012imagenet}. Besides, we reimplement two other approaches. One is a semi-supervised learning method called Pseudo-Label~\cite{lee2013pseudo}, which exploits classifier's prediction as ground truth for unlabeled data. The other one is the Bottom-Up method introduced in~\cite{sukhbaatar2014learning}, where the relation between noisy labels and clean labels are built by a confusion matrix $Q$. In the experiments, we directly use the true $Q$ as shown in Fig.~\ref{fig:cm_strong_weak} instead of estimating its values.

We list all the experiment settings in Tab.~\ref{tab:exp}. Different methods require different training data. We use only the clean data $\mathcal{D}_c$ to get the baselines under strong supervisions. On the other hand, when all the data are used, we upsample the clean data as discussed in Sec.~\ref{sec:deep_learning_from_noisy_labels}. Meanwhile, the noisy labels of $\mathcal{D}_{\eta}$ are treated as true labels for AlexNet, and are discarded for Pseudo-Label.

In general, we use a mini-batch size of $256$. The learning rate is initialized to be $0.001$ and is divided by $10$ after every $50,000$ iterations. We keep training each model until convergence. Classification accuracies on the test set are presented in Tab.~\ref{tab:exp}.

@@

\subsection{Evaluation on CIFAR-10 with Synthetic Noises} % (fold)
\label{sub:evaluation_on_synthetic_dataset}
We also conduct synthetic experiments on CIFAR-10 following the settings of~\cite{sukhbaatar2014learning}. We first randomly generate a confusion matrix $Q$ between clean labels and noisy labels, and then corrupt the training labels according to it. Based on Caffe's CIFAR10-quick model, we compare the~\cite{sukhbaatar2014learning} (Bottom Up with true $Q$) with our model under different noise levels. The test accuracies are reported in Tab.~\ref{tab:accuracy_synthetic}.

@@

\subsection{Effect of Noise Estimation} % (fold)
\label{sub:effect_of_noise_estimation}
In order to understand how our model handles noisy labels, we first show several examples in Fig.~\ref{fig:model_output}. We find that given a noisy label, our model exploits its current knowledge to estimate the probability distribution of the true label and then replaces the noisy one with it as supervision. Another interesting observation is that our model can still figure out the correct label if the prediction of the class label $p(\vec{y}|\vec{x})$ or the noise type $p(\vec{z}|\vec{x})$ goes wrong. These two latent variables compensate with each other to help the system distinguish between hard samples and noisy labeled ones.

Next we demonstrate the effectiveness of learning to predict the label noise type. Notice that if an image has low probability of ``noise free'' (\ie, $p(\vec{z}_1=1|\vec{x})$ is small), then our model will believe it is likely to be mislabeled. In order to check the reliability of these predictions, we first estimate $p(\vec{z}_1=1|\vec{x})$ on the validation set and sort the images accordingly in ascending order. Then for each image, we check whether its noisy label mismatches its clean label. The rank-precision curve is plotted in Fig.~\ref{fig:noise_prediction_rp}. It shows that the intuitive observation --- mislabeled images often share similar patterns --- is reasonable, and our model is trained to recognize such patterns to identify potentially mislabeled images.

@@

\section{Conclusion} % (fold)
\label{sec:conclusion}
In this paper, we address the problem of training a deep learning classifier with massive amount of noisy labeled training data and a small amount of clean annotations which are generally easy to collect. To utilize both limited clean labels and massive noisy labels, we propose a novel probabilistic framework to describe how noisy labels are produced. We introduce two latent variables, the true label and the noise type, to bridge the gap between an input image and its noisy label. We observe that the label noises not only depend on the ambiguity between classes, but could follow similar visual patterns as well. We build the dependency of the noise type \wrt Images, and infer the ground truth label with EM algorithm. We integrate the probabilistic model into a deep learning framework and demonstrate the effectiveness of our method on a large-scale real-world clothing classification dataset.
% section conclusion (end)

>>>>

Our goal is to build an end-to-end deep learning system that is capable of training with both limited clean labels and massive noisy labels more effectively. Figure~\ref{fig:overview} shows the framework of our approach. We collect $1,000,000$ clothing images from online shopping websites. Each image is automatically assigned with a noisy label according to the keywords in its surrounding text. We manually refine $72,409$ image labels, which constitute a clean sub-dataset. All the data are then used to train CNNs, while the major challenge is to identify and correct wrong labels during the training process.

To cope with this challenge, we extend CNNs with a novel probabilistic model, which infers the true labels and uses them to supervise the training of the network. Our work is inspired by~\cite{sukhbaatar2014learning}, which modified a CNN by inserting a linear layer on top of the softmax layer to map clean labels to noisy labels. However,~\cite{sukhbaatar2014learning} assumed noisy labels are conditionally independent with input images given clean labels. By examining our collected dataset, we find that this assumption is too strong to fit real-world data well. For example, in Figure~\ref{fig:noise_cluster}, all the images should belong to ``Hoodie''. The top five are correct while the bottom five are either mislabeled as ``Windbreaker'' or ``Jacket''. Since different sellers have their own bias on different categories, they may provide wrong keywords for similar clothes. We can capture these visual patterns to estimate how likely an image is mislabeled. Based on these observations, we further introduce two types of label noise:
\begin{itemize}
\item \textbf{Confusing noise} makes the noisy label reasonably wrong. It usually occurs when the image content is confusing (\eg, the samples with ``?'' in Figure~\ref{fig:overview}).
\item \textbf{Pure random noise} makes the noisy label totally wrong. It is often caused by either the mismatch between an image and its surrounding text, or the false conversion from the text to label (\eg, the samples with ``$\times$'' in Figure~\ref{fig:overview}).
\end{itemize}

@@

\section{Label Noise Model} % (fold)
\label{sec:label_noise_model}
We target on learning a classifier from a set of images with noisy labels. To be specific, we have a noisy labeled dataset $\mathcal{D}_{\eta}=\left\{\left(\vec{x}^{(1)}, \tilde{y}^{(1)}\right), \dots, \left(\vec{x}^{(N)}, \tilde{y}^{(N)}\right)\right\}$ with $n$-th image $\vec{x}^{(n)}$ and its corresponding noisy label $\tilde{y}^{(n)} \in \{1,\dots,L\}$, where $L$ is the number of classes. We describe how the noisy label is generated by using a probabilistic graphical model shown in Figure~\ref{fig:noise_pgm}.

@@

Following this assignment rule, we define the conditional probability of the noisy label as
\begin{equation} \label{eq:likelihood}
    p(\tilde{\vec{y}}|\vec{y},\vec{z}) = \begin{cases}
        \tilde{\vec{y}}^{T}\mat{I}\vec{y} & \text{if } \vec{z}_1 = 1 \\
        \frac{1}{L-1}\tilde{\vec{y}}^{T}(\mat{U} - \mat{I})\vec{y} & \text{if } \vec{z}_2 = 1 \\
        \tilde{\vec{y}}^{T}\mat{C}\vec{y} & \text{if } \vec{z}_3 = 1,
    \end{cases}
\end{equation}
where $\mat{I}$ is the identity matrix, $\mat{U}$ is the unit matrix (all the elements are ones), $\mat{C}$ is a sparse stochastic matrix with $\mathrm{tr}(\mat{C})=0$ and $\mat{C}_{ij}$ denotes the confusion probability between classes $i$ and $j$. Then we can derive from Figure~\ref{fig:noise_pgm} the joint distribution of $\tilde{\vec{y}}, \vec{y}$ and $\vec{z}$ conditioning on $\vec{x}$,
\begin{equation} \label{eq:joint_distribution}
  p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}) = p(\tilde{\vec{y}} | \vec{y}, \vec{z}) p(\vec{y} | \vec{x}) p(\vec{z} | \vec{x}).
\end{equation}

While the class label probability distribution $p(\vec{y} | \vec{x})$ is comprehensible, the semantic meaning of $p(\vec{z} | \vec{x})$ needs extra clarification: it represents how confusing the image content is. Specific to our clothing classification problem, $p(\vec{z} | \vec{x})$ can be affected by different factors, including background clutter, image resolution, the style and material of the clothes. Some examples are shown in Figure~\ref{fig:compare_noise_level}.

@@

We integrate the proposed label noise model into a deep learning framework. As demonstrated in Figure~\ref{fig:framework}, we predict the probability $p(\vec{y}|\vec{x})$ and $p(\vec{z}|\vec{x})$ by using two independent CNNs. Moreover, we append a label noise model layer at the end, which takes as input the CNNs' prediction scores and the observed noisy label. Stochastic Gradient Ascent with backpropagation technique is used to approximately optimize the whole network. In each forward pass, the label noise model layer computes the posterior of latent variables according to Eq.~\eqref{eq:posterior}. While in the backward pass, it computes the gradients according to Eq.~\eqref{eq:gradient}.

@@

After both CNN components are properly pretrained, we can start to train the whole network with massive noisy labeled data. However, some practical issues need further discussion. First, if we merely use noisy labels, we will lose precious knowledge that we have gained before and the model could be drifted. Therefore, we need to mix the data with clean labels into our training set, which is depicted in Figure~\ref{fig:framework} as the extra supervisions for the two CNNs. Then each CNN receives two kinds of gradients, one is from the clean labels and the other is from the noisy labels. We denote them by $\Delta_c$ and $\Delta_n$, respectively. A potential problem is that $|\Delta_c| \ll |\Delta_n|$, because clean data is much less than noisy data. To deal with this problem, we bootstrap the clean data $\mathcal{D}_c$ to half amount of the noisy data $\mathcal{D}_{\eta}$. In our experiments, we find that the performance of the classifier drops significantly without upsampling, but it is not sensitive with the upsampling ratio as long as the number of clean and noisy samples are in the same order.

@@

The confusion matrix between clean and noisy labels is presented in Figure~\ref{fig:cm_strong_weak}. We can see that the overall accuracy is $61.54\%$, and some pairs of classes are very confusing with each other (\eg Knitwear and Sweater), which means that the noisy labels are not so reliable.

@@

\subsection{Evaluation on the Collected Dataset} % (fold)
\label{sub:evaluation}
We validate our method through a series of experiments conducted on the collected dataset. Our implementation is based on Caffe~\cite{jia2014caffe}, and the \texttt{bvlc\_reference\_caffenet}\footnote{\url{http://caffe.berkeleyvision.org/model_zoo.html}} is chosen as the baseline model, which approximates AlexNet~\cite{krizhevsky2012imagenet}. Besides, we reimplement two other approaches. One is a semi-supervised learning method called Pseudo-Label~\cite{lee2013pseudo}, which exploits classifier's prediction as ground truth for unlabeled data. The other one is the Bottom-Up method introduced in~\cite{sukhbaatar2014learning}, where the relation between noisy labels and clean labels are built by a confusion matrix $Q$. In the experiments, we directly use the true $Q$ as shown in Figure~\ref{fig:cm_strong_weak} instead of estimating its values.

We list all the experiment settings in Table~\ref{tab:exp}. Different methods require different training data. We use only the clean data $\mathcal{D}_c$ to get the baselines under strong supervisions. On the other hand, when all the data are used, we upsample the clean data as discussed in Section~\ref{sec:deep_learning_from_noisy_labels}. Meanwhile, the noisy labels of $\mathcal{D}_{\eta}$ are treated as true labels for AlexNet, and are discarded for Pseudo-Label.

In general, we use a mini-batch size of $256$. The learning rate is initialized to be $0.001$ and is divided by $10$ after every $50,000$ iterations. We keep training each model until convergence. Classification accuracies on the test set are presented in Table~\ref{tab:exp}.

@@

\subsection{Evaluation on CIFAR-10 with Synthetic Noises} % (fold)
\label{sub:evaluation_on_synthetic_dataset}
We also conduct synthetic experiments on CIFAR-10 following the settings of~\cite{sukhbaatar2014learning}. We first randomly generate a confusion matrix $Q$ between clean labels and noisy labels, and then corrupt the training labels according to it. Based on Caffe's CIFAR10-quick model, we compare the~\cite{sukhbaatar2014learning} (Bottom Up with true $Q$) with our model under different noise levels. The test accuracies are reported in Table~\ref{tab:accuracy_synthetic}.

@@

\subsection{Effect of Noise Estimation} % (fold)
\label{sub:effect_of_noise_estimation}
In order to understand how our model handles noisy labels, we first show several examples in Figure~\ref{fig:model_output}. We find that given a noisy label, our model exploits its current knowledge to estimate the probability distribution of the true label and then replaces the noisy one with it as supervision. Another interesting observation is that our model can still figure out the correct label if the prediction of the class label $p(\vec{y}|\vec{x})$ or the noise type $p(\vec{z}|\vec{x})$ goes wrong. These two latent variables compensate with each other to help the system distinguish between hard samples and noisy labeled ones.

Next we demonstrate the effectiveness of learning to predict the label noise type. Notice that if an image has low probability of ``noise free'' (\ie, $p(\vec{z}_1=1|\vec{x})$ is small), then our model will believe it is likely to be mislabeled. In order to check the reliability of these predictions, we first estimate $p(\vec{z}_1=1|\vec{x})$ on the validation set and sort the images accordingly in ascending order. Then for each image, we check whether its noisy label mismatches its clean label. The rank-precision curve is plotted in Figure~\ref{fig:noise_prediction_rp}. It shows that our model can identify mislabeled samples quite well based on their image content, which verifies the observation that mislabeled images often share similar visual patterns.

@@

\section{Conclusion} % (fold)
\label{sec:conclusion}
In this paper, we address the problem of training a deep learning classifier with a massive amount of noisy labeled training data and a small amount of clean annotations which are generally easy to collect. To utilize both limited clean labels and massive noisy labels, we propose a novel probabilistic framework to describe how noisy labels are produced. We introduce two latent variables, the true label and the noise type, to bridge the gap between an input image and its noisy label. We observe that the label noises not only depend on the ambiguity between classes, but could follow similar visual patterns as well. We build the dependency of the noise type \wrt images, and infer the ground truth label with the EM algorithm. We integrate the probabilistic model into a deep learning framework and demonstrate the effectiveness of our method on a large-scale real-world clothing classification dataset.
% section conclusion (end)
